Convergence rate of MAP estimates \\ for the exponential family

### 2021-09-27 main.tex
Detail the step-size with weird exponent

\subsection{Other Idea}
There are two main bottlenecks when casting MAP as an iterative algorithm. 
\begin{enumerate}
	\item The first one is the variance definition which always end up being infinite when asked to be uniform on every possible trajectories, if the log-partition  has some barrier structure.
	\item The second one is the step-size scheduling. $O(\inv{t})$ is decreasing too fast for most proof techniques. Even in the Euclidean setting where the variance does not matter, it results in a disappointing rate $O(\inv{\log t})$.
\end{enumerate}
Regarding this second point, it might be possible to do something comparable to \citet{gower2019sgd} who get a $O(\inv{t})$ rate with a $O(\inv{t})$ step-size. Take $m$ and $L$ as the (relative) strong-convexity and smoothness constants of the problem.Their descent lemma looks like
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^2 \sigma^2
\end{align}
where $B_t$ is some measure of suboptimality, in this case the $\ell^2$ distance between $\natp^t$ and $\natp^*$ ; and $\lr_t \leq \inv{2L}$.
On the other hand, the problem of most Bregman analysis is that they do not get this $\lr^2$ in front the variance. Instead they simply get a $\lr$, because they do not benefit from the homogeneity of the norm
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t \sigma^2 \; .
\end{align}
The problem is that this exponent $2$ is crucial to get fast convergence with a $O(\inv{t})$ step-size. Otherwise the variance is too large and we need to decrease the step-size more slowly to reach the optimum quickly.
Now our question is what happens if we get something in-between ? e.g.
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^\alpha \sigma^2 \; .
\end{align}
where $1<\alpha<2$ is some exponent characterizing the geometry of the problem.
Inspired by \citet{gower2019sgd} we set
\begin{align}
	\lr_t := \inv{m} \left ( 1 - \left (\frac{k}{k+1} \right)^\alpha \right ) \; .
\end{align}

>>>>

\subsection{Other Idea}
There are two main bottlenecks when casting MAP as an iterative algorithm. 
\begin{enumerate}
	\item The first one is the variance definition which always end up being infinite when asked to be uniform on every possible trajectories, if the log-partition  has some barrier structure.
	\item The second one is the step-size scheduling. $O(\inv{t})$ is decreasing too fast for most proof techniques. Even in the Euclidean setting where the variance does not matter, it results in a disappointing rate $O(\inv{\log t})$.
\end{enumerate}
Regarding this second point, it might be possible to do something comparable to \citet{gower2019sgd} who get a $O(\inv{t})$ rate with a $O(\inv{t})$ step-size. Take $m$ and $L$ as the (relative) strong-convexity and smoothness constants of the problem, along with $K=\frac{L}{m}$ the condition number.Their descent lemma looks like
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^2 \sigma^2
\end{align}
where $B_t$ is some measure of suboptimality, in this case the $\ell^2$ distance between $\natp^t$ and $\natp^*$ ; and $\lr_t \leq \inv{2L}$.
On the other hand, the problem of most Bregman analysis is that they do not get this $\lr^2$ in front the variance. Instead they simply get a $\lr$, because they do not benefit from the homogeneity of the norm
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t \sigma^2 \; .
\end{align}
The problem is that this exponent $2$ is crucial to get fast convergence with a $O(\inv{t})$ step-size. Otherwise the variance is too large and we need to decrease the step-size more slowly to reach the optimum quickly.
Now our question is what happens if we get something in-between ? e.g.
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^\alpha \sigma^2 \; .
\end{align}
where $1<\alpha<2$ is some exponent characterizing the geometry of the problem.
Inspired by \citet{gower2019sgd} we set
\begin{align}
	\lr_t := \inv{m} \left ( 1 - \left (\frac{t}{t+1} \right)^\alpha \right ) \; .
\end{align}
Asymptotically, we are in the desired category
\begin{align}
	\lr_t := \inv{m} \left ( 1 - \left (1 - \frac{1}{t+1} \right)^\alpha \right ) 
	\approx \inv{m} \left ( 1 - \left (1 - \frac{\alpha}{t+1} \right) \right ) 
	\approx \frac{\alpha}{m t} \in O(\inv{t} \; .
\end{align}
When does it start to $\lr_t \leq \inv{2L}$ ?
\begin{align}
	\lr_t \leq \inv{2L}
	&\iff 1 - \left (\frac{t}{t+1} \right)^\alpha \leq \inv{2K} \\
	&\iff \left (1 - \inv{2K}  \right)^{\inv{\alpha}} \leq \frac{t}{t+1}\\
	&\iff \frac{t+1}{t} \leq \left (1 - \inv{2K}  \right)^{-\inv{\alpha}}
\end{align}
and we know by convexity of $(1+x)^\beta$ that $	\left (1 - \inv{2K}  \right)^{-\inv{\alpha}} \geq 1 + \inv{2\alpha K}$.
This is giving us a sufficient condition on $t$ :
\begin{align}
	\frac{t+1}{t}  \leq 1 + \inv{2\alpha K} 
	&\iff t ( 1 - 1 -  \inv{2\alpha K}) + 1 \leq 0 \\
	&\iff t \geq 2\alpha K \; .
\end{align}
Thus when $t < t_0 = \ceil{2\alpha K}$, we set $\lr_t = \inv{2L}$.

\TODO{Questions for later : finish typing the proof of convergence. Then think about how this exponent could appear?}
